---
title             : "Children spontaneously re-create core properties of language in a new modality"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["library.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library(papaja)
library(tidyverse)
library(ggthemes)
library(langcog)
library(lme4)
library(exactRankTests)
```

@bruner1974communication


# General setup and procedure

All studies used the same general setup and procedure. We established an audio-video connection between two separate rooms within a child laboratory. In each room a Panasonic XX video camera was mounted on a tripod which was placed behind a XX" XXX TV screen. Each camera captured an area in front of the TV, which was marked with black tape on the floor. Children were encouraged to stay within the marked area to ensure that their partner could see them. A black cross in the center of the area marked children's starting position in each trial. A little fence was used in each room to prevent participants from approaching the TV and/or the camera. The connection was established through DVI cables that went from the camera in one room to the TV in the other room and vice versa. External microphones were plugged into the cameras to capture children's speech. The direct connection between camera and TV made sure that that participants could interact in a smooth and contingent way. One of the rooms was assigned to be the Production Room (PR) and the other the Comprehension Room (CR). A third camera was put in the back of the CR, capturing the child as well as the TV screen. This was an additional measure to ensure that we could link the two videos recorded by the two main cameras.

The two rooms were located on the same hallway. When stepping outside, the experimenters (E1 and E2) could easily talk to one another and coordinate their actions. In each room, we installed a picture board. Importantly, the picture boards were never visible on the TV in the other room. In part 1, the picture board was a wooden, turn-able wheel with five triangular pictures, showing 1) a hammering hand, 2) a female person combing her hair, 3) a boy on a bicycle, 4) a woman eating with a fork and 5) nothing (a white piece of paper, see Figure X). On top of the wheel was a little red arrow that was used to indicate which picture children had to communicate to their partner. A picture was selected by turning the wheel until the picture was on top underneath the arrow. In part 2, the board was a larger, rectangular board. The pictures differed from experiment to experiment and are therefore described later in more detail. In part 2, a sticky red paper arrow was used to indicate a picture.

In the beginning of each trial, E1 was with the child in PR (Production Child - PC) and E2 with the child in CR (Comprehension Child - CC). The general trial structure in all experiments went as follows: Children were first asked to take their starting position (facing away from the board towards the wall). Next, E1 selected a picture in PR. On a visual signal from E1, both Es said "ready, set, go" and left the room. Children were now allowed to turn around. The task of the child in PR was to communicate the content of the selected picture to the child in CR. The child in CR made a choice by selecting a picture on her board and knocking on the door. Thereupon, Es opened the doors, E2 told E1 across the hallway which picture had been selected in the CR and E1 checked if this was the same picture as in the PR. If the pictures were different, children were told that it didn't work out this time but that they could try again in the next trial. If the pictures were the same, children received a marble, which they collected in a tube. At the end of the study, children exchanged the marbles for stickers. 

All children started with a training phase in which the video as well as the audio connection was still intact. In the beginning of the training phase, children and Es entered their respective rooms. They greeted each other through the TV and established that they could only be seen by their partner when they were inside the marked area. Next, E1 and PC "discovered" the picture board. E1 encouraged PC to ask CC if they also had a picture board in their room. When CC confirmed that they did, PC went through all pictures (including the empty one) and asked CC if had the same kind of picture on her board. This established common ground between children that they had the same pictures on their board. Es did not provide label for the pictures but let children label them. Es used whatever label children used to refer to the pictures. Next, E1 suggested that they could play a fun game in which children could win marbles. Children were told that the game involves PC communicating to CC which picture was selected on her board and CC selecting the same picture. This marked the beginning of the training phase in which the audio connection was still working. PC could therefore simply tell CC which picture was selected. The training familiarized children with the structure of the game. During training, each picture was selected once. 

After the last training trial, children were told that they could play the game in a different way to make it more fun. Children and Es left the room and met in the hallway. Next, one E went into each room and cut the audio connection by turning off the volume at each TV. To strengthen the impression that the other one could not longer hear what was being said as well as to prevent children from trying to communicate via shouting, low instrumental Jazz music was played in both rooms (Chet Baker - "Out of nowhere"). If children shouted nevertheless, the experimenters asked them to lower their voice. When children re-entered the room, each E drew her child's attention to the music and encouraged her to ask the other child whether she also heard the music. This way children discovered that they could no longer hear each other. After stating that they could not hear anybody from the other room, children and Es left the room to meet in the hallway. There they confirmed that they could not hear each other when in the room. E1 asked the children if they wanted to play the game nevertheless and also reminded them of the game's structure. Gestures as an alternative means of communication was never mentioned. Next, children went into their respective room and the first test trial started.

In each test trial, children had one minute to establish communication. If they remained passive, E1 entered the PR and prompted PC to use iconic gestures in three steps. Between prompts, she waited for 30s to see if the child would start gesturing. As first prompt, E1 said:"[CCs name] can't hear you, but she can see you, maybe you can *show* (German: zeigen) it to her". The second prompt was: "She can't hear you but she can see you, look we can wave to her, maybe you can *demonstrate* (German: vormachen) it to her". As the final prompt, she said: "Oh, I have an idea ... this way it could work" and started to perform an iconic gesture depicting the action corresponding to the selected picture. As soon as the child also used the gesture, E1 left the room and the trial continued as described above. During prompting, E2 entered CR and encouraged CC to attend to the TV.

In the following, we describe the methodological details and deviations from the general protocoll for each study.

# Part 1.1.1: Spontaneous production of gestures by 4- and 6-year olds in peer context 

## Participants
We tested 24 same sex, same age dyads of four- and six-year-olds (12 dyads per age group, six male and six female). Children came from an ethnically homogeneous, mid-sized German city (approx. 550.000 inhabitants, median income €1254 per month as of 2015), were mostly mono-lingual and had mixed socio-economic background. They were recruited from a database of children whose parents volunteered to take part in studies on child development. Data collection took place between April and September 2016. XX additional dyads were excluded because ...

## Design and procedure
The test phase following the training was divided in two parts, "spontaneous" and "uptake". Children were first tested with all pictures except the empty picture. We reasoned that including the empty picture from the beginning might discourage children from establishing communiation through gestures. The empty picture was introduced after each other picture had been selected at least once. The first trial of spontaneous assessed whether children would discover the idea to use gestures to communicate independently. If the PR remained passive for one minute in the first test trial, instead of receiving prompts, children switched rooms and the former CR got a chance to spontaneously start gesturing. If the new PR remained passive for another minute, E1 started prompting. The three trials follwoing the first trial, each with a new picture, assessed whether children would transfer the idea of using gestures. For the second part, uptake, children switched rooms and roles. In this part, all pictures were used from the beginning.

Six-year-olds received a total of 30 test trials, 20 in spontaneous and 10 in uptake. Pilot testing suggested that this was too much for the younger age group and so four-year-olds received a total of 19 test trials, 14 in spontaneous and 5 in uptake. The order in which pictures were selected was randomized with the constraint that each picture was selected once before another picture could be selected again. 

The intermediate prompts were not helpful for most children and we therefore do not differentiate between them in the analysis. We only analyse whether a dyad received a prompt or not.

## Results and discussion

```{r part 1-1-1, echo=FALSE}
d1 <- read.csv(file="data/pt1_prod_up_comp.csv") %>% 
  filter(block == "Spontaneous") %>%
  na.omit(prompts) # remove trials without 


# Do 4yo need more prompts in trial 1 than 6yo?
d1.t <- d1 %>%  
  filter(blocktrial == 1) %>%
  group_by(age, dyad_id) %>%
  summarise(prompts = sum(prompts)) %>%
  #mutate(prompts = ifelse(prompts > 0,1,0)) %>%
  summarize(prompts = list(prompts)) %>%
  spread(age,prompts) %>%
  mutate(sum4 = sum(unlist(M48)),
         sum6 = sum(unlist(M72)),
         stat = wilcox.exact(unlist(M48),unlist(M72))$statistic,
         p_value = wilcox.exact(unlist(M48),unlist(M72))$p.value)%>%
  select(sum4,sum6,stat,p_value) %>%
  knitr::kable(digits = 5)

# Do they get better with trials? Are 6yo better than 4yo?
d1.m <- d1 %>%
  #filter(age=="M48") %>%
  filter(blocktrial < 5) %>%
  mutate(blocktrial = scale(as.numeric(blocktrial), center = TRUE, scale=TRUE))%>%
  group_by(blocktrial,dyad_id,age) %>%
  summarise(prompts = sum(prompts))%>%
  mutate(prompts = ifelse(prompts > 0,1,0))
  
  
m1 <- glmer(prompts ~ blocktrial+age + (blocktrial|dyad_id), family = binomial, data = d1.m, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m1)
drop1(m1 ,test="Chi")

# plot for dyads producing gestures (instead of for dyads needing prompts)
d1.p <- d1 %>%  
  mutate(prompts = ifelse(prompts > 0,1,0)) %>%
  mutate(Age = ifelse(age == "M48","4yo","6yo")) %>%
  filter(blocktrial<5) %>%
  group_by(Age, blocktrial) %>%
  summarise(prompts = sum(prompts)) %>%
  mutate(prompts = 12 - prompts)


ggplot(d1.p, 
       aes(x = blocktrial, y = prompts, col = Age)) +
  geom_line(aes(group= Age))+
  geom_point(aes(group= Age))+
  theme_few() + 
  scale_y_continuous(name = "Dyads Spontaneously Producing Gestures", limits = c(0,12), breaks = c(0,2,4,6,8,10,12))+
  scale_x_continuous(name = "Trial", breaks = c(1:4))+
  scale_colour_solarized()



```


# Part 1.1.2: Spontaneous production of gestures by 3- and 4-year-olds when communicationg with an adult.

## Participants

## Design and procedure

## Results and discussion
```{r part 1-1-2}
d2 <- read.csv(file="data/pt1_trans_3_4.csv") 

# do 4yo transfer more rapidly compared to 3yo?
# 3-way interaction: slope for trial is steeper in 4yo only in transfer but not in imitation

d2.m <- d2 %>%
  mutate(trial = scale(as.numeric(blocktrial), center = TRUE, scale=TRUE))

m2 <- glmer(prompts ~ block*age*blocktrial + (blocktrial + block||id) + (1|stimulus) , family = binomial, data = d2.m,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m2)

drop1(m2,test="Chi")

# plot 
d2.p <- d2 %>%
  mutate(Age = ifelse(age == "m36","3yo","4yo")) %>%
  mutate(block = relevel(block, ref = "Transfer")) %>%
  group_by(Age,trial, blocktrial,block) %>%
  summarise(prompts = sum(prompts))%>%
  mutate(prompts = 12 - prompts)

ggplot(d2.p, 
       aes(x = blocktrial, y = prompts, col = Age)) +
  geom_line(aes(group= Age))+
  geom_point(aes(group= Age))+
  facet_grid(~block)+
  theme_few() + 
  scale_y_continuous(name = "Dyads Spontaneously Producing Gestures", limits = c(0,12), breaks = c(0,2,4,6,8,10,12))+
  scale_x_continuous(name = "Trial", breaks = c(1:6))+
  scale_colour_solarized()

```


# Part 1.2: Uptake

## Participants

## Design and procedure

## Results and discussion
```{r part 1-2}
# prompts needed during uptake phase
d3 <- read.csv(file="data/pt1_prod_up_comp.csv") %>% 
  filter(block == "Uptake") %>%
  na.omit(prompts)

d3 %>%  
  filter(blocktrial == 1) %>%
  group_by(age, dyad_id) %>%
  summarise(prompts = sum(prompts)) %>%
  mutate(prompts = ifelse(prompts > 0,1,0)) %>%
  summarize(prompts = list(prompts)) %>%
  spread(age,prompts) %>%
  mutate(sum4 = sum(unlist(M48)),
         sum6 = sum(unlist(M72)),
         stat = wilcox.exact(unlist(M48),unlist(M72))$statistic,
         p_value = wilcox.exact(unlist(M48),unlist(M72))$p.value)%>%
  select(sum4,sum6,stat,p_value) %>%
  knitr::kable(digits = 5)

  
# plot
d3.p <- d3 %>%  
  mutate(prompts = ifelse(prompts > 0,1,0)) %>%
  mutate(Age = ifelse(age == "M48","4yo","6yo")) %>%
  filter(trial<5) %>%
  group_by(Age, trial) %>%
  summarise(prompts = sum(prompts))%>%
  mutate(prompts = 12 - prompts)

# plot
ggplot(d3.p, 
       aes(x = trial, y = prompts, col = Age)) +
  geom_line(aes(group= Age))+
  geom_point(aes(group= Age))+
  theme_few() + 
  scale_y_continuous(name = "Dyads Spontaneously Producing Gestures", limits = c(0,12), breaks = c(0,2,4,6,8,10,12))+
  scale_x_continuous(name = "Trial", breaks = c(1:4))+
  scale_colour_solarized()



```

# Part 1.3: Comprehension

## Participants

## Design and Procedure

## Results and discussion
```{r pt 1-3}
# comprehension by phase excluding empty picture

d4 <- read.csv(file="data/pt1_prod_up_comp.csv") %>%
  mutate(Age = ifelse(age == "M48","4yo","6yo")) %>%
  na.omit(comp)

# Comparing comprehension against chance (25% correct)
d4 %>%
  group_by(Age, block, dyad_id)%>%
  summarise(comp = mean(comp))%>%
  summarise(comp = list(comp)) %>%
  group_by(Age, block) %>%
  mutate(df= t.test(unlist(comp), mu = 0.25)$parameter,
         t_value = t.test(unlist(comp), mu = 0.25)$statistic,
         p_value = t.test(unlist(comp), mu = 0.25)$p.value) %>%
  select(Age, block ,df,t_value,p_value) %>%
  knitr::kable(digits = 5)


# are 6yo better than 4yo at comprehension

d4.m <- d4 %>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=TRUE))

m4 <- glmer(comp ~ block+age + (block|dyad_id), family = binomial, data = d4.m,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m4)

drop1(m4,test="Chi")


# plot per age group and phase
d4.p <- d4 %>%
  group_by(Age, block, dyad_id)%>%
  summarise(comp = mean(comp))

d4.p1 <- d4.p %>%
  multi_boot_standard(col = "comp")

ggplot() +
  geom_jitter(data = d4.p, aes(x = block, y = comp, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d4.p1,aes(x = block, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .5), size = 1) + 
  geom_hline(yintercept = 0.25, lty=2)+
  guides(alpha = F)+
  #facet_grid(~block)+
  scale_y_continuous(name = "Proportion Correct", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()
```


# Part 1.4: Abstract concepts (empty picture)

## Participants

## Design and Procedure

## Results and discussion
```{r part 1-4}
#### Production
# prompts per age group and phase
d5 <- read.csv(file="data/pt1_prod_up_comp.csv") %>% 
  filter(stimulus == "Nothing") %>%
  mutate(Age = ifelse(age == "M48","4yo","6yo")) %>%
  na.omit(prompts)

# descriptives for production (plot not informative) and tests
# do 4yo need more prompts compared to 6yo?
d5 %>%
  group_by(age, block, dyad_id) %>%
  summarise(prompts = sum(prompts)) %>%
  mutate(prompts = ifelse(prompts > 0,1,0)) %>%
  summarize(prompts = list(prompts)) %>%
  spread(age,prompts) %>%
  group_by(block) %>%
  mutate(sum4 = sum(unlist(M48)),
         sum6 = sum(unlist(M72)),
         stat = wilcox.exact(unlist(M48),unlist(M72))$statistic,
         p_value = wilcox.exact(unlist(M48),unlist(M72))$p.value)%>%
  select(block,sum4,sum6,stat,p_value) %>%
  knitr::kable(digits = 5)

### Comprehension 
#comparing against chance per age group and phase
d5 %>%
  group_by(Age, block, dyad_id)%>%
  summarise(comp = mean(comp))%>%
  summarise(comp = list(comp)) %>%
  group_by(Age,block) %>%
  mutate(mean = mean(unlist(comp)),
         df= t.test(unlist(comp), mu = 0.25)$parameter,
         t_value = t.test(unlist(comp), mu = 0.25)$statistic,
         p_value = t.test(unlist(comp), mu = 0.25)$p.value) %>%
  select(Age,block,mean,df,t_value,p_value) %>%
  knitr::kable(digits = 3)

# are 6yo better than 4yo at comprehension

d5.m <- d5 %>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=TRUE))

m5 <- glmer(comp ~ block+age + (1|dyad_id), family = binomial, data = d5.m,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m5)

drop1(m5,test="Chi")

# plot per age group and phase
d5.p2 <- d5 %>%
  group_by(Age, block, dyad_id)%>%
  summarise(comp = mean(comp))

d5.p3 <- d5.p2 %>%
  multi_boot_standard(col = "comp")

ggplot() +
  geom_jitter(data = d5.p2, aes(x = block, y = comp, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d5.p3,aes(x = block, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .5), size = 1) + 
  geom_hline(yintercept = 0.25, lty=2)+
  guides(alpha = F)+
  #facet_grid(~block)+
  scale_y_continuous(name = "Proportion Correct", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()


```



# Part 1.5: Convergence

## Participants

## Design and Procedure

## Results and discussion
```{r part 1-5}
d6 <- read.csv(file="data/pt1_convergence.csv") %>% 
  mutate(Age = ifelse(age == "M48","4yo","6yo"))

# Do raters choose the gesture from the same dyad above chance in both age groups?
d6 %>%
  group_by(Age, rater)%>%
  summarise(corr = mean(corr))%>%
  summarise(corr = list(corr)) %>%
  group_by(Age) %>%
  mutate(mean = mean(unlist(corr)),
         df= t.test(unlist(corr), mu = 0.5)$parameter,
         t_value = t.test(unlist(corr), mu = 0.5)$statistic,
         p_value = t.test(unlist(corr), mu = 0.5)$p.value) %>%
  select(Age,mean,df,t_value,p_value) %>%
  knitr::kable(digits = 3)

# plot per age group
d6.p <- d6 %>%
  group_by(Age, rater)%>%
  summarise(corr = mean(corr))

d6.p1 <- d6.p %>%
  multi_boot_standard(col = "corr")

ggplot() +
  geom_jitter(data = d6.p, aes(x = 1, y = corr, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d6.p1,aes(x = 1, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  #facet_grid(~block)+
  guides(alpha = F)+
  theme_few() + 
  scale_y_continuous(name = "Proportion Gesture from Same Dyad Chosen", limits = c(0,1))+
  scale_colour_solarized()+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.title.x=element_blank())
  

```


# Part 1.6: Drift to arbitrary

## Production

### Participants

### Design and Procedure

### Results and discussion
```{r part 1-6 production}
d7 <- read.csv(file="data/pt1_drift_production.csv") %>% 
  mutate(Age = ifelse(age == "M48","4yo","6yo"))%>%
  filter(phase != "Uptake")  #filtering out uptake because 4yo only got 1 trial per stimulus in uptake


# Do raters choose the later gesture as the more abstract one?
### t-tests against chance per age group and day
d7 %>%
  group_by(Age, day, rater)%>%
  summarise(corr = mean(corr))%>%
  summarise(corr = list(corr)) %>%
  group_by(Age, day) %>%
  mutate(mean = mean(unlist(corr)),
         df= t.test(unlist(corr), mu = 0.5)$parameter,
         t_value = t.test(unlist(corr), mu = 0.5)$statistic,
         p_value = t.test(unlist(corr), mu = 0.5)$p.value) %>%
  select(Age,day,mean,df,t_value,p_value) %>%
  knitr::kable(digits = 3)


### better with more repititions? That is, 6yo better than 4yo and day 2 better than day 1 (not the best model since pictures on day 2 were different compared to day 1).

m7 <- glmer(corr ~ Age+day+(Age+day||rater)+(1|stimulus), family = binomial, data = d7, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(m7)

drop1(m7,test="Chi")


# plot per age group  and day
d7.p <- d7 %>%
  group_by(Age, day, rater)%>%
  summarise(corr = mean(corr))

d7.p1 <- d7.p %>%
  multi_boot_standard(col = "corr")

ggplot() +
  geom_jitter(data = d7.p, aes(x = day, y = corr, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d7.p1,aes(x = day, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  #facet_grid(~day,scales = "free_x", space = "free_x")+
  guides(alpha = F)+
  scale_y_continuous(name = "Proportion Later Gesture Chosen", limits = c(0,1))+
  theme_few() + 
  scale_colour_solarized()

```


## Comprehension
```{r part 1-6 comprehension}
d8 <- read.csv(file="data/pt1_drift_comprehension.csv") %>% 
  mutate(Age = ifelse(age == "M48","4yo","6yo"))

# are children better in test compared to pretest when seeing abstract gestures?
d8.m <- d8  %>%
  filter(condition == "Abstract") %>%
  mutate(trial = scale(as.numeric(trial), center = TRUE, scale=TRUE))

# model 
m8 <- glmer(corr ~ age+phase + (trial+phase|id) + (age+phase|stimulus), family = binomial, data = d8.m, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
# summary
summary(m8)
# p values
drop1(m8, test = "Chi")


### t-tests against chance per age group, day and phase
d8 %>%
  group_by(Age,condition, phase, id)%>%
  summarise(corr = mean(corr))%>%
  summarise(corr = list(corr)) %>%
  group_by(Age,condition, phase) %>%
  mutate(mean = mean(unlist(corr)),
         df= t.test(unlist(corr), mu = 1/3)$parameter,
         t_value = t.test(unlist(corr), mu = 1/3)$statistic,
         p_value = t.test(unlist(corr), mu = 1/3)$p.value) %>%
  select(Age,phase,condition,mean,df,t_value,p_value) %>%
  knitr::kable(digits = 3)


# plot per age group, condition and phase
d8.p <- d8 %>%
  group_by(Age, condition, phase,id)%>%
  summarise(corr = mean(corr))

d8.p1 <- d8.p %>%
  multi_boot_standard(col = "corr")

ggplot() +
  geom_jitter(data = d8.p, aes(x = condition, y = corr, col = Age, alpha = .2), width = .1, height = .05)+
  geom_pointrange(data = d8.p1,aes(x = condition, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1) + 
  geom_hline(yintercept = 1/3, lty=2)+
  facet_grid(~phase)+
  guides(alpha = F)+
  scale_y_continuous(name = "Proportion Later Chosen", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()
```
### Participants

### Design and Procedure

### Results and discussion

# Part 1 Appendix: Plots by stimulus

## Comprehension
```{r plots by stimulus: comprehension}
# plot by stimulus, age group and phase
d4.p2 <- d4 %>%
  group_by(Age, stimulus, block, dyad_id)%>%
  summarise(comp = mean(comp))

d4.p3 <- d4.p2 %>%
  multi_boot_standard(col = "comp")


d4.p4 <- d4 %>%
  group_by(Age, block, dyad_id)%>%
  summarise(comp = mean(comp)) %>%
  multi_boot_standard(col = "comp")%>%
  mutate(stimulus = "Overall")

ggplot() +
  geom_jitter(data = d4.p2, aes(x = stimulus, y = comp, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d4.p3,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .5), size = 0.5, alpha = .5) +
  geom_pointrange(data = d4.p4,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .5), size = 1) +
  geom_hline(yintercept = 0.25, lty=2)+
  guides(alpha = F)+
  facet_grid(~block)+
  scale_y_continuous(name = "Proportion Correct", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()+ coord_flip()
```

## Convergence
```{r plots by stimulus: convergence}

# plot per age group and stimulus
d6.p2 <- d6 %>%
  group_by(Age, stimulus, rater)%>%
  summarise(corr = mean(corr))

d6.p3 <- d6.p2 %>%
  multi_boot_standard(col = "corr")


d6.p4 <- d6 %>%
  group_by(Age,rater) %>%
  summarise(corr = mean(corr)) %>%
  multi_boot_standard(col = "corr") %>%
  mutate(stimulus = "Overall")


ggplot() +
  geom_jitter(data = d6.p2, aes(x = stimulus, y = corr, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d6.p3,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1, alpha = .5) + 
  geom_pointrange(data = d6.p4,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  #facet_grid(~block)+
  guides(alpha = F)+
  scale_y_continuous(name = "Proportion Gesture from Same Dyad Chosen", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()+ coord_flip()

```

## Drift to arbitrary
```{r plots by stimulus: drift}
# plot per age group, phase, day and stimulus
d7.p2 <- d7 %>%
  group_by(Age, day, stimulus, rater)%>%
  summarise(corr = mean(corr))

d7.p3 <- d7.p2 %>%
  multi_boot_standard(col = "corr")


d7.p4 <- d7 %>%
  group_by(Age,day,rater) %>%
  summarise(corr = mean(corr)) %>%
  multi_boot_standard(col = "corr") %>%
  mutate(stimulus = "Overall")

ggplot() +
  geom_jitter(data = d7.p2, aes(x = stimulus, y = corr, col = Age, alpha = .2), width = .1,height = .05)+
  geom_pointrange(data = d7.p3,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1, alpha = .5) +
  geom_pointrange(data = d7.p4,aes(x = stimulus, y = mean, col = Age,ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = .5), size = 1) +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_vline(xintercept = 8.5, lty=1)+
  facet_grid(day~.,scales = "free_y", space = "free_y", drop = T)+
  guides(alpha = F)+
  scale_y_continuous(name = "Proportion Later Chosen", limits = c(-0.05,1.05))+
  theme_few() + 
  scale_colour_solarized()+ 
  coord_flip()
```
# Part 2.1: Size

## Participants

## Design and Procedure

## Results and discussion


# Part 2.2: Number

## Participants

## Design and Procedure

## Results and discussion


# Part 2.3: Movement

## Participants

## Design and Procedure

## Results and discussion


# Part 2.4: Transitive actions with two agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2.5: Transitive actions with three agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2 overall analysis

## Participants

## Design and Procedure

## Results and discussion



\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
