---
title             : "Children spontaneously re-create core properties of language in a new modality"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["library.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library(papaja)
```

@bruner1974communication

# General setup and procedure

Hallo Manuel,
danke f?r den Code von gestern. Ich kann jetzt ?nderungen, die ich lokal vorgenommen habe in das online repository pushen. Habe das jetzt auch drei mal gemacht und f?r den wichtigsten Git Code mal ein R-file angelegt mit den wichtigsten Befehlen f?r uns. 

Hi Gregor,

das scheint zu funktionieren. Ich hab einfach git pull verwendet und somit mein file auf den neusten stand gebracht. Alles wunderbar. 
Ich hab ein paar kommentare in dem git commands file gemacht. Kannst du ja mal ausprobieren.
Ich push das alles gleich nochmal und füge auch einen Ordner in dem wir pdfs sammeln können. 




# Part 1.1: Spontaneous production of gestures

## Participants

## Design and Procedure

## Results and discussion


# Part 1.2: Uptake

## Participants

## Design and Procedure

## Results and discussion


# Part 1.3: Comprehension

## Participants

## Design and Procedure

## Results and discussion


# Part 1.4: Abstract concepts (empty picture)

## Participants

## Design and Procedure

## Results and discussion


# Part 1.5: Convergence

## Participants

## Design and Procedure

## Results and discussion


# Part 1.6: Drift to arbitrary

## Production

### Participants

### Design and Procedure

### Results and discussion

## Comprehension

### Participants

### Design and Procedure

### Results and discussion


# Part 2.1: Size

## Participants

## Design and Procedure

## Results and discussion


# Part 2.2: Number

## Participants

## Design and Procedure

## Results and discussion


# Part 2.3: Movement

## Participants

## Design and Procedure

## Results and discussion


# Part 2.4: Transitive actions with two agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2.5: Transitive actions with three agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2 overall analysis

## Participants

## Design and Procedure

## Results and discussion



\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
