---
title             : "Children spontaneously re-create core properties of language in a new modality"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["library.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library(papaja)
```

@bruner1974communication

# General setup and procedure

Hallo Manuel,
danke für den Code von gestern. Ich kann jetzt Änderungen, die ich lokal vorgenommen habe in das online repository pushen. Habe das jetzt auch drei mal gemacht und für den wichtigsten Git Code mal ein R-file angelegt mit den wichtigsten Befehlen für uns. 


# Part 1.1: Spontaneous production of gestures

## Participants

## Design and Procedure

## Results and discussion


# Part 1.2: Uptake

## Participants

## Design and Procedure

## Results and discussion


# Part 1.3: Comprehension

## Participants

## Design and Procedure

## Results and discussion


# Part 1.4: Abstract concepts (empty picture)

## Participants

## Design and Procedure

## Results and discussion


# Part 1.5: Convergence

## Participants

## Design and Procedure

## Results and discussion


# Part 1.6: Drift to arbitrary

## Production

### Participants

### Design and Procedure

### Results and discussion

## Comprehension

### Participants

### Design and Procedure

### Results and discussion


# Part 2.1: Size

## Participants

## Design and Procedure

## Results and discussion


# Part 2.2: Number

## Participants

## Design and Procedure

## Results and discussion


# Part 2.3: Movement

## Participants

## Design and Procedure

## Results and discussion


# Part 2.4: Transitive actions with two agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2.5: Transitive actions with three agents 

## Participants

## Design and Procedure

## Results and discussion


# Part 2 overall analysis

## Participants

## Design and Procedure

## Results and discussion



\newpage

# References
```{r create_r-references}
r_refs(file = "library.bib")
```

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
